---
title: "Comp 4442 Final Project"
author: "Luke Sonnanburg"
output: word_document
---

```{r setup, include=FALSE}
#install.packages('fastDummies')
#install.packages('ggcorrplot')
#install.packages(c("FactoMineR", "factoextra"))
#install.packages('GPArotation')
library(GPArotation)
library("FactoMineR")
library("factoextra")
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(gapminder)
library(leaps)
library(glmnet)
library(ggplot2)
library(MASS)
library(usdm)
library(car)
library(corrplot)
library(texreg)
library(MPV)
library(caret)
library(glmnet)
library(psych)
library(tidyverse)
library(broom)
library(fastDummies)
library(ggcorrplot)
library(glmnet)

```

## Project Overview
```{r, include= FALSE}
# Importing Data
math<-read.csv("student-mat.csv", sep = ';') # Student data and scores in math
port<-read.csv("student-por.csv", sep = ';') # Student data and scors in portuguese 

```

```{r}
math
```

```{r}
math$avg <- (math$G1 + math$G2 + math$G3)/3
math.grades <- math[,c(31:34)]
boxplot(math.grades)

```
Getting data into the correct form
```{r}
math$school <- as.numeric(ifelse(math$school =="GP", 0, 1)) # Gabriel Pereira = 0, Mousinho da Silveira = 1
math$sex <- as.numeric(ifelse(math$sex == "F", 0, 1)) # Female = 0, Male = 1
math$address <- as.numeric(ifelse(math$address == "U", 0, 1)) # Urban = 0, Rural = 1
math$famsize <- as.numeric(ifelse(math$famsize == "LE3", 0, 1)) # Less than 3 kids = 0, more = 1
math$Pstatus <- as.numeric(ifelse(math$Pstatus == "T", 0, 1)) # Parents together = 0, parents apart = 1
math$Medu <- as.numeric(ifelse(math$Medu == 4, 1,0))
math$Fedu <- as.numeric(ifelse(math$Fedu == 4, 1,0))
math$guardian <- as.numeric(ifelse(math$guardian == "other", 0, 1)) # recoding legal guardian as factors
math$reason <- ifelse(math$reason %in% c('reputation','course'), 1,0 ) # 1 -> chose school for academic reasons
math$schoolsup <- as.numeric(ifelse(math$schoolsup =="no", 0, 1))
math$famsup <- as.numeric(ifelse(math$famsup =="no", 0, 1))
math$paid <- as.numeric(ifelse(math$paid =="no", 0, 1))
math$activities <- as.numeric(ifelse(math$activities =="no", 0, 1))
math$nursery <- as.numeric(ifelse(math$nursery =="no", 0, 1))
math$higher <- as.numeric(ifelse(math$higher =="no", 0, 1))
math$internet <- as.numeric(ifelse(math$internet =="no", 0, 1))
math$romantic <- as.numeric(ifelse(math$romantic =="no", 0, 1))
math <- math[-c(9:10,31:33)]
mathX <- math[-c(29)]
mathY <- math$avg
```

The variable inflation factor is a measure of excessive multicollinearity that may be a cause for
concern. The values here are small enough not to worry about.
```{r}
naive.model <- lm(avg~., data = math)
vif(naive.model)
par(mfrow = c(2, 2))
plot(naive.model)
```

```{r fig.height = 8, fig.width = 8}
math.cor <- cor(math)

ggcorrplot(math.cor, title = 'Student Data Correlation Plot')
```



```{r}
KMO(math.cor)
```



The skree plot shows eigenvalues of principal factors. Greater eigenvalues indicate that a factor
explains more of the output variable's variance. 
Generally eigenvalues >= 1 imply the factor explains more variance than a single variable would.
In this case, the first two factors have eigenvalues >=1 and should obviously be included.
A more subjective means of determining how many factors is where the scree plot levels off.
In this case, the first place where the eigenvalues flatten (if only for a moment) is at 
the fourth factor. Since adding factors until the next "leveling off" would result in 8
factors where more are uninfluential than influential, I'll operate on the assumption of four
factors being reasonable.

(https://www.theanalysisfactor.com/factor-analysis-1-introduction/)

```{r}
parallel <- fa.parallel(mathX, fa='fa')
parallel
```
While the scree plot for portuguese test data suggests 5 factors may be more appropriate,
the ideal scenario would be if both data sets produced similar factor loadings to paint
a picture of what a "successful" student looks like. For portuguese I'll proceed with 4 to 
match math, but also experiment with 5.


```{r}
math.varimaxfit.4 <- fa(r=mathX, nfactors = 4, rotate="varimax",fm="pa")

names(math.varimaxfit.4$loadings) <- c('a','b','c','d') 
math.varimaxfit.4
```
```{r fig.height = 10, fig.width = 10}
fa.diagram(math.varimaxfit.4)
```



```{r}
math.varimaxfit.4$loadings

```


```{r}
math.varimaxfit.5 <- fa(r=mathX, nfactors = 5, rotate="varimax",fm="pa")
math.varimaxfit.5
```
```{r fig.height = 10, fig.width = 10}
math.varimaxfit.4 <- fa(r=mathX, nfactors = 4, rotate="varimax",fm="pa")
math.varimaxfit.4
fa.diagram(math.varimaxfit.4)
```







```{r fig.height = 10, fig.width = 10}
colnames(math.varimaxfit.4$loadings) <- c('Outgoing/Social', 'Academic Drive/Background', 'Hardship', 'Rural Location')
math.varimaxfit.4$loadings
fa.diagram(math.varimaxfit.4)
```

```{r}
math.dat <- data.frame(math.varimaxfit.4$scores, math$avg)
names(math.dat) <- c('Outgoing/Social', 'Academic Drive/Background', 'Hardship', 'Rural Location', 'Scores')
math.dat
```
